---
title: "Beta Coefficient Plotting (EDA)"
output:
  pdf_document: default
  html_notebook: default
---

# Front Matter

```{r results = FALSE, include = FALSE}
# libraries
library(maps)
library(sf)
library(data.table)
library(tidyverse)
library(haven)
library(RColorBrewer)
library(classInt)
library(mosaic)
library(sp)
```

```{r}
# read data and shape files
### map shape file
ukroblast <-
  st_read("ukraineGIS/gadm36_UKR_1.shp")
### regional gdp data
gdpraw <- 
  fread("grossregionalproduct.csv")
### exit poll 1999
exit99 <-
  read_sav("KIIS/Exit-poll1999/Exit-poll1999_1_eng.sav")
### exit poll 2002
exit02 <-
  read_sav("KIIS/Exit-poll2002/Exitpoll2002_unicode.sav")
### expectations 2002
expect <-
  read_sav("KIIS/Expectations2002/SOC_EXP_massive.sav")
### gl2000
gl2000 <-
  read_sav("KIIS/GL2000/KIIS_public_opinion2000_ENG.sav")
### national integrity 1998
integrity <-
  read_sav("KIIS/Integrity1998/KIIS_PSP_1998_UKR_unicode.sav")
### omnibus nov 2002
omn_nov02 <-
  read_sav("KIIS/KIIS_OMN2002_nov/KIIS_omn2002nov_ENG.sav")
### omnibus feb 2002
omn_feb02 <-
  read_sav("KIIS/Politika2002_feb2/KIIS-KCPIK_politika2002-feb2_RUS_unicode.sav")
### views politika 2002
views02 <-
  read_sav("KIIS/Politika2002-feb1/KIIS-KCPIK_politika2002-feb1_RUS_unicode.sav")
### usia1997
usia97 <-
  read_sav("KIIS/USIA1997/USIA1997_ENG.sav")
### unicef
unicef <-
  read_sav("UNICEF/hh.sav")
### CIDI data
cidi_raw <-
  fread("ukrmain_RU_061419.csv")
### population data
pop2004_raw <-
  fread("pop_2004.csv")
```

# Map Exploration

```{r}
### look at names
ukroblast$NAME_1
```

# Data Aggregation

```{r}
# clean up pop data
pop2004 <-
  pop2004_raw %>%
  mutate(Oblast = str_replace_all(Oblast, "[^[:alnum:]]", " "),
         Oblast = ifelse(Oblast == "Autonomous Republic of Crimea", "Crimea", 
                  ifelse(Oblast == "Dnipropetrovsk", "Dnipropetrovs'k",
                  ifelse(Oblast == "Donetsk", "Donets'k",
                  ifelse(Oblast == "Ivano Frankivsk", "Ivano-Frankivs'k",
                  ifelse(Oblast == "Khmelnytskiy", "Khmel'nyts'kyy",
                  ifelse(Oblast == "City of Kyiv", "Kiev City",
                  ifelse(Oblast == "Kyiv", "Kiev",
                  ifelse(Oblast == "Lviv", "L'viv",
                  ifelse(Oblast == "Luhansk", "Luhans'k",
                  ifelse(Oblast == "Odesa", "Odessa",
                  ifelse(Oblast == "City of Sevastopol", "Sevastopol'",
                  ifelse(Oblast == "Ternopil", "Ternopil'",
                  ifelse(Oblast == "Zakarpattya", "Transcarpathia", Oblast))))))))))))))

# clean up GDP data, obtain variables of interest
gdp <-
  gdpraw[6:32] %>%
  select(V1, V29, V2)
colnames(gdp) <-
  c("Oblast", "PerCapitaGDP", "GDP")
gdp1 <-
  gdp %>%
  mutate(PerCapitaGDP = as.numeric(PerCapitaGDP),
         GDP = 1000000 * as.numeric(GDP), # GDP in dataset by millions
         Oblast = ifelse(Oblast == "Autonomous Republic of Crimea", "Crimea", 
                  ifelse(Oblast == "Dnipropetrovsk", "Dnipropetrovs'k",
                  ifelse(Oblast == "Donetsk", "Donets'k",
                  ifelse(Oblast == "Ivano-Frankivsk", "Ivano-Frankivs'k",
                  ifelse(Oblast == "Khmelnytskiy", "Khmel'nyts'kyy",
                  ifelse(Oblast == "Kyiv" & PerCapitaGDP == 23130, "Kiev City",
                  ifelse(Oblast == "Kyiv" & PerCapitaGDP == 6652, "Kiev",
                  ifelse(Oblast == "Lviv", "L'viv",
                  ifelse(Oblast == "Luhansk", "Luhans'k",
                  ifelse(Oblast == "Odesa", "Odessa",
                  ifelse(Oblast == "Sevastopol", "Sevastopol'",
                  ifelse(Oblast == "Ternopil", "Ternopil'",
                  ifelse(Oblast == "Zakarpattya", "Transcarpathia", Oblast)))))))))))))) %>%
  left_join(pop2004, by = "Oblast")
gdp1 <-
  gdp1 %>%
  mutate(PerCapitaGDP = ifelse(Oblast == "Crimea", (gdp1$GDP[1] + gdp1$GDP[27])/(gdp1$Population[1] + gdp1$Population[27]),
                        ifelse(Oblast == "Sevastopol'", (gdp1$GDP[1] + gdp1$GDP[27])/(gdp1$Population[1] + gdp1$Population[27]),
                               PerCapitaGDP)))
  

# how widespread is corruption?
### taking proportion of people who answered 4: everyone is corrupt
corruptionraw <-
  table(gl2000$V237, gl2000$V45) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # calcuates percent of people answering everyone is corrupt
  mutate(all_corrupt = `4`/Sum) %>% ## there is a problem here because the 27th row is not sevastopol, it is the total
  rownames_to_column()
corruption <-
  corruptionraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         all_corrupt = ifelse(Oblast == "Sevastopol'", all_corrupt[1], all_corrupt)) %>%
  arrange(Oblast) %>%
  select(Oblast, all_corrupt) 


# deric I'm not sure why you're doing this part
# corruption$all_corrupt[20]<-
#  corruption$all_corrupt[4]

# what language do you speak at home?
### percent Ukrainian
languageraw <-
  table(gl2000$V237, gl2000$V108) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # calculate percent of Ukrainian speakers
  mutate(ukr = `1`/Sum) %>% #check data here, are there really 0 ukrainian speakers in Crimea?
  rownames_to_column()

language <-
  languageraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         ukr = ifelse(Oblast == "Sevastopol'", ukr[1], ukr)) %>%
  arrange(Oblast) %>%
  select(Oblast, ukr)

# for which party or block have you voted?
### percent "Our Ukraine"
politicalraw <-
  table(exit02$V4, exit02$V6) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # percent "Our Ukraine" party
  mutate(our_ukr = `2`/Sum) %>%
  rownames_to_column()

# cleaning political raw data
political <-
  politicalraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         our_ukr = ifelse(Oblast == "Sevastopol'", our_ukr[1], our_ukr)) %>%
  arrange(Oblast) %>%
  select(Oblast, our_ukr) 

# is the necessary medical service available?
### percent "not at all sure"
medicalraw <-
  table(expect$v121, expect$v48) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # percent "not at all sure"
  mutate(med_no = `5`/Sum) %>%
  rownames_to_column()
medical <-
  medicalraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         med_no = ifelse(Oblast == "Sevastopol'", med_no[1], med_no)) %>%
  arrange(Oblast) %>%
  select(Oblast, med_no)

# get used to 10 year changes?
### percent "certainly succeeded"
tenyearraw <-
  table(omn_feb02$V64, omn_feb02$V5) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # percent "certainly succeeded"
  mutate(certainly10 = `1`/Sum) %>%
  rownames_to_column()
tenyear <-
  tenyearraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         certainly10 = ifelse(Oblast == "Sevastopol'", certainly10[1], certainly10)) %>%
  arrange(Oblast) %>%
  select(Oblast, certainly10)

# religiosity
### percent "very important"
religionraw <-
  table(usia97$V175, usia97$V154) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # percent "very important"
  mutate(rel_veryimp = `1`/Sum) %>%
  rownames_to_column()
religion <-
  religionraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         rel_veryimp = ifelse(Oblast == "Sevastopol'", rel_veryimp[1], rel_veryimp)) %>%
  arrange(Oblast) %>%
  select(Oblast, rel_veryimp)

# economic attitude, oblast
### how would you assess the economic situation in your city? percent saying very bad
econraw <-
  table(gl2000$V237, gl2000$V113) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # percent "very bad"
  mutate(econ_bad = `5`/Sum) %>%
  rownames_to_column()
econ <-
  econraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         econ_bad = ifelse(Oblast == "Sevastopol'", econ_bad[1], econ_bad)) %>%
  arrange(Oblast) %>%
  select(Oblast, econ_bad)

# environmental concerns, oblast
### percent concerned about air pollution
environmentraw <-
  table(usia97$V175, usia97$V107) %>%
  addmargins() %>%
  as.data.frame.matrix() %>%
  # calculate percent
  mutate(air_concern = `2`/Sum) %>%
  rownames_to_column()
environment <-
  environmentraw %>%
  mutate(Oblast = derivedVariable("Crimea" = rowname == 1,
                                "Kiev City" = rowname == 2,
                                "Kiev" = rowname == 3,
                                "Vinnytsya" = rowname == 4,
                                "Volyn" = rowname == 5,
                                "Dnipropetrovs'k" = rowname == 6,
                                "Donets'k" = rowname == 7,
                                "Zhytomyr" = rowname == 8,
                                "Transcarpathia" = rowname == 9,
                                "Zaporizhzhya" = rowname == 10,
                                "Ivano-Frankivs'k" = rowname == 11,
                                "Kirovohrad" = rowname == 12,
                                "Luhans'k" = rowname == 13,
                                "L'viv" = rowname == 14,
                                "Mykolayiv" = rowname == 15,
                                "Odessa" = rowname == 16,
                                "Poltava" = rowname == 17,
                                "Rivne" = rowname == 18,
                                "Sumy" = rowname == 19,
                                "Ternopil'" = rowname == 20,
                                "Kharkiv" = rowname == 21,
                                "Kherson" = rowname == 22,
                                "Khmel'nyts'kyy" = rowname == 23,
                                "Cherkasy" = rowname == 24,
                                "Chernihiv" = rowname == 25,
                                "Chernivtsi" = rowname == 26,
                                "Sevastopol'" = rowname == 27),
         # fix sevastopol to take on crimea values since row 27 is avg across ukraine
         air_concern = ifelse(Oblast == "Sevastopol'", air_concern[1], air_concern)) %>%
  arrange(Oblast) %>%
  select(Oblast, air_concern)


# CIDI data clean up and calculating proportions by each oblast
cidi <-
  cidi_raw %>%
  filter(!is.na(de40)) %>%
  
  # mental conditions
  mutate(anx = ifelse(DSM_AGO == 1|DSM_SO == 1|DSM_GAD == 1|DSM_PDS == 1, 1, 0),
         ied = ifelse(DSM_IEDH == 1, 1, 0),
         aff = ifelse(DSM_MDE == 1|DSM_DYS == 1, 1, 0),
         alc = ifelse(DSM_ALA == 1|DSM_ALD == 1, 1, 0),
         any = ifelse(anx == 1|ied == 1|aff == 1|alc == 1, 1, 0),
         Oblast = derivedVariable("Crimea" = io15 == 1,
                                "Kiev City" = io15 == 2,
                                "Kiev" = io15 == 3,
                                "Vinnytsya" = io15 == 4,
                                "Volyn" = io15 == 5,
                                "Dnipropetrovs'k" = io15 == 6,
                                "Donets'k" = io15 == 7,
                                "Zhytomyr" = io15 == 8,
                                "Transcarpathia" = io15 == 9,
                                "Zaporizhzhya" = io15 == 10,
                                "Ivano-Frankivs'k" = io15 == 11,
                                "Kirovohrad" = io15 == 12,
                                "Luhans'k" = io15 == 13,
                                "L'viv" = io15 == 14,
                                "Mykolayiv" = io15 == 15,
                                "Odessa" = io15 == 16,
                                "Poltava" = io15 == 17,
                                "Rivne" = io15 == 18,
                                "Sumy" = io15 == 19,
                                "Ternopil'" = io15 == 20,
                                "Kharkiv" = io15 == 21,
                                "Kherson" = io15 == 22,
                                "Khmel'nyts'kyy" = io15 == 23,
                                "Cherkasy" = io15 == 24,
                                "Chernihiv" = io15 == 25,
                                "Chernivtsi" = io15 == 26),
         x = 1) %>% #placeholder to help add up all of the people in that Oblast
  group_by(Oblast) %>%
  summarize(totalpop = sum(x),
            anx = sum(anx),
            ied = sum(ied),
            aff = sum(aff),
            alc = sum(alc),
            any = sum(any)) %>%
  select(Oblast, anx, ied, aff, alc, any, totalpop) %>%
  mutate(anx_prop = anx/totalpop,
         ied_prop = ied/totalpop,
         aff_prop = aff/totalpop,
         alc_prop = alc/totalpop,
         any_prop = any/totalpop)


# Create one large data set by joining with oblasts - case file readout
cidi_final <- 
  rbind(cidi, cidi[rep(4, 1),])
cidi_final[27,]$Oblast <- "Sevastopol'"
combined_final <-
  cidi_final %>%
  left_join(corruption, by = "Oblast") %>%
  left_join(econ, by = "Oblast") %>%
  left_join(environment, by = "Oblast") %>%
  left_join(gdp1, by = "Oblast") %>%
  left_join(language, by = "Oblast") %>%
  left_join(medical, by = "Oblast") %>%
  left_join(political, by = "Oblast") %>%
  left_join(religion, by = "Oblast") %>%
  left_join(tenyear, by = "Oblast") %>%
  select(-totalpop, -anx_prop, -ied_prop, -aff_prop, -alc_prop, -any_prop, -Population) %>%
  mutate(Oblast = str_replace_all(Oblast, "[^[:alnum:]]", ""))

fwrite(combined_final, "case_file.csv")

```


### Barrington's 8 Region Map

```{r}
# categorize oblasts into Barrington's 8 regions
ukroblast$region8 <-
  derivedFactor(East = ukroblast$NAME_1 == "Donets'k"|ukroblast$NAME_1 == "Luhans'k",
                Eastcentral = ukroblast$NAME_1 == "Kharkiv"|ukroblast$NAME_1 == "Zaporizhzhya"|
                  ukroblast$NAME_1 == "Dnipropetrovs'k",
                Krym = ukroblast$NAME_1 == "Crimea"|ukroblast$NAME_1 == "Sevastopol'",
                South = ukroblast$NAME_1 == "Kherson"|ukroblast$NAME_1 == "Mykolayiv"|ukroblast$NAME_1 == "Odessa",
                Northcentral = ukroblast$NAME_1 == "Chernihiv"|ukroblast$NAME_1 == "Sumy"|
                  ukroblast$NAME_1 == "Poltava"|ukroblast$NAME_1 == "Cherkasy"|ukroblast$NAME_1 == "Kirovohrad"|
                  ukroblast$NAME_1 == "Kiev"|ukroblast$NAME_1 == "Kiev City",
                Westcentral = ukroblast$NAME_1 == "Khmel'nyts'kyy"|ukroblast$NAME_1 == "Zhytomyr"|
                  ukroblast$NAME_1 == "Vinnytsya"|ukroblast$NAME_1 == "Rivne"|ukroblast$NAME_1 == "Volyn",
                West = ukroblast$NAME_1 == "Ivano-Frankivs'k"|ukroblast$NAME_1 == "L'viv"|ukroblast$NAME_1 == "Ternopil'",
                Southwest = ukroblast$NAME_1 == "Chernivtsi"|ukroblast$NAME_1 == "Transcarpathia")

# plot data
ggplot(data = ukroblast, aes(fill = region8)) +
  geom_sf() +
  labs(title = "Barrington's 8 Region Classification",
       fill = "Regions") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())
```

### GDP Per Capita (2004)

```{r}
# find bins, visualize
quantile(gdp1$PerCapitaGDP)
ggplot(gdp1, aes(x = PerCapitaGDP)) +
  geom_histogram() +
  geom_vline(xintercept = 4735.5) +
  geom_vline(xintercept = 5163.0) +
  geom_vline(xintercept = 6840.0)

# bin GDP per capita data, add to shape file
gdp2 <-
  gdp1 %>%
  mutate(GDPcapita = derivedFactor("< 4740" = PerCapitaGDP < 4740, 
                                   "4740-5160" = PerCapitaGDP >= 4740 & PerCapitaGDP < 5160,
                                   "5160-6840" = PerCapitaGDP >= 5160 & PerCapitaGDP < 6840, 
                                   "> 6840" = PerCapitaGDP >= 6840)) %>%
  select(Oblast, GDPcapita) %>%
  arrange(Oblast)
ukroblast$percapitaGDP <-
  gdp2$GDPcapita
remove(gdp2)

# plot
ggplot(data = ukroblast, aes(fill = percapitaGDP)) +
  geom_sf() +
  labs(title = "Per Capita GDP",
       fill = "GDP per Capita") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())


# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on gdp


```

### Proportion who think everyone is corrupt

```{r}
# find bins, visualize
quantile(corruption$all_corrupt)
ggplot(corruption, aes(x = all_corrupt)) +
  geom_histogram() +
  geom_vline(xintercept = 0.3222) +
  geom_vline(xintercept = 0.3846) +
  geom_vline(xintercept = 0.5172)

# bin data, add to shape file
corruption2 <-
  corruption %>%
  mutate(corrupt_bin = derivedFactor("< 0.32" = all_corrupt < 0.32,
                                     "0.32-0.38" = all_corrupt >= 0.32 & all_corrupt < 0.38,
                                     "0.38-0.51" = all_corrupt >= 0.38 & all_corrupt < 0.51,
                                     "> 0.51" = all_corrupt >= 0.51))
ukroblast$all_corrupt <-
  corruption2$corrupt_bin
remove(corruption2)

# plot
ggplot(data = ukroblast, aes(fill = all_corrupt)) +
  geom_sf() +
  labs(title = "Proportion who think everyone is corrupt",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on corruption

```


### Proportion speaking Ukrainian at home

```{r}
# find bins, visualize
quantile(language$ukr)
ggplot(language, aes(x = ukr)) +
  geom_histogram() +
  geom_vline(xintercept = 0.2870) +
  geom_vline(xintercept = 0.6111) +
  geom_vline(xintercept = 0.8882)

# bin data, add to shape file
language2 <-
  language %>%
  mutate(ukr_bin = derivedFactor("< 0.29" = ukr < 0.29,
                                 "0.29-0.61" = ukr >= 0.29 & ukr < 0.61,
                                 "0.61-0.88" = ukr >= 0.61 & ukr < 0.88,
                                 "> 0.88" = ukr >= 0.88))
ukroblast$ukrlang <-
  language2$ukr_bin
remove(language2)

# plot
ggplot(data = ukroblast, aes(fill = ukrlang)) +
  geom_sf() +
  labs(title = "Proportion who speak Ukrainian at home",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on proportion who speak ukr at home

```

### Voting Patterns - Proportion voting for "Our Ukraine"

```{r}
# find bins, visualize
quantile(political$our_ukr)
ggplot(political, aes(x = our_ukr)) +
  geom_histogram() +
  geom_vline(xintercept = 0.0813) +
  geom_vline(xintercept = 0.25) +
  geom_vline(xintercept = 0.3986)

# bin data, add to shape file
political2 <-
  political %>%
  mutate(ukr_bin = derivedFactor("< 0.09" = our_ukr < 0.09,
                                 "0.09-0.25" = our_ukr >= 0.09 & our_ukr < 0.25,
                                 "0.25-0.40" = our_ukr >= 0.25 & our_ukr < 0.40,
                                 "> 0.40" = our_ukr >= 0.40))
ukroblast$our_ukr <-
  political2$ukr_bin
remove(political2)

# plot
ggplot(data = ukroblast, aes(fill = our_ukr)) +
  geom_sf() +
  labs(title = "Proportion voting for 'Our Ukraine' Party",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on pol party
```

### Is necessary medical service availabe? Proportion who say no

```{r}
# find bins, visualize
quantile(medical$med_no)
ggplot(medical, aes(x = med_no)) +
  geom_histogram() +
  geom_vline(xintercept = 0.1914) +
  geom_vline(xintercept = 0.2295) +
  geom_vline(xintercept = 0.3361)

# bin data, add to shape file
medical2 <-
  medical %>%
  mutate(bin = derivedFactor("< 0.20" = med_no < 0.20,
                             "0.20-0.23" = med_no >= 0.20 & med_no < 0.23,
                             "0.23-0.34" = med_no >= 0.23 & med_no < 0.34,
                             "> 0.34" = med_no >= 0.34))
ukroblast$med_no <-
  medical2$bin
remove(medical2)

# plot
ggplot(data = ukroblast, aes(fill = med_no)) +
  geom_sf() +
  labs(title = "Proportion who think necessary medical service is not available",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on med service availability

```

### How successful were you to adapting to changes in the past 10 year? Proportion - certainly successful

```{r}
# find bins, visualize
quantile(tenyear$certainly10)
ggplot(tenyear, aes(x = certainly10)) +
  geom_histogram() +
  geom_vline(xintercept = 0.0356) +
  geom_vline(xintercept = 0.0846) +
  geom_vline(xintercept = 0.1061)

# bin data, add to shape file
tenyear2 <-
  tenyear %>%
  mutate(bin = derivedFactor("< 0.04" = certainly10 < 0.04,
                             "0.04-0.08" = certainly10 >= 0.04 & certainly10 < 0.08,
                             "0.08-0.10" = certainly10 >= 0.08 & certainly10 < 0.1,
                             "> 0.10" = certainly10 >= 0.1))
ukroblast$certainly10 <-
  tenyear2$bin
remove(tenyear2)

# plot
ggplot(data = ukroblast, aes(fill = certainly10)) +
  geom_sf() +
  labs(title = "Proportion who were certainly successful in adapting to 10 year changes",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on adaptability

```

### Proportion saying religion is very important

```{r}
# find bins, visualize
quantile(religion$rel_veryimp)
ggplot(religion, aes(x = rel_veryimp)) +
  geom_histogram() +
  geom_vline(xintercept = 0.1262) +
  geom_vline(xintercept = 0.2188) +
  geom_vline(xintercept = 0.3117)

# bin data, add to shape file
religion2 <-
  religion %>%
  mutate(bin = derivedFactor("< 0.13" = rel_veryimp < 0.13,
                             "0.13-0.22" = rel_veryimp >= 0.13 & rel_veryimp < 0.22,
                             "0.22-0.31" = rel_veryimp >= 0.22 & rel_veryimp < 0.31,
                             "> 0.31" = rel_veryimp >= 0.31))
ukroblast$rel_veryimp <-
  religion2$bin
remove(religion2)

# plot
ggplot(data = ukroblast, aes(fill = rel_veryimp)) +
  geom_sf() +
  labs(title = "Proportion who think religion is very important",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on importance of religion

```

### Economic Attitude - Proportion thinking economic situation in their city is very bad

```{r}
# find bins, visualize
quantile(econ$econ_bad)
ggplot(econ, aes(x = econ_bad)) +
  geom_histogram() +
  geom_vline(xintercept = 0.3096) +
  geom_vline(xintercept = 0.4222) +
  geom_vline(xintercept = 0.5759)

# bin data, add to shape file
econ2 <-
  econ %>%
  mutate(bin = derivedFactor("< 0.31" = econ_bad < 0.31,
                             "0.31-0.42" = econ_bad >= 0.31 & econ_bad < 0.42,
                             "0.42-0.58" = econ_bad >= 0.42 & econ_bad < 0.58,
                             "> 0.58" = econ_bad >= 0.58))
ukroblast$econ_bad <-
  econ2$bin
remove(econ2)

# plot
ggplot(data = ukroblast, aes(fill = econ_bad)) +
  geom_sf() +
  labs(title = "Proportion thinking economic situation in their city is very bad",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on econ attitudes

```

### Environmental - Proportion concerned about air pollution

```{r}
# find bins, visualize
quantile(environment$air_concern)
ggplot(environment, aes(x = air_concern)) +
  geom_histogram() +
  geom_vline(xintercept = 0.1183) +
  geom_vline(xintercept = 0.1667) +
  geom_vline(xintercept = 0.2501)

# bin data, add to shape file
environment2 <-
  environment %>%
  mutate(bin = derivedFactor("< 0.12" = air_concern < 0.12,
                             "0.12-0.17" = air_concern >= 0.12 & air_concern < 0.17,
                             "0.17-0.25" = air_concern >= 0.17 & air_concern < 0.25,
                             "> 0.25" = air_concern >= 0.25))
ukroblast$air_concern <-
  environment2$bin
remove(environment2)

# plot
ggplot(data = ukroblast, aes(fill = air_concern)) +
  geom_sf() +
  labs(title = "Proportion who are concerned about air pollution",
       fill = "Proportion") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank())

# TODO: beta coefficient calculation for each oblast regressing disorder prevalence on air pollution

```
















