---
title: "Frequency Tables"
author: "Deric Liang"
date: "28 June 2019"
output:
  pdf_document: default
  html_notebook: default
---

# Front Matter

```{r include = FALSE, results='hide'}
# libraries
library(tidyverse)
library(survey)
library(data.table)
library(mosaic)
```

```{r}
# read data
cidi_raw <-
  fread("ukrmain_RU_061419.csv")
```

```{r}
# REPRODUCING THE FOUR CORE CONDITIONS; REGION VARIABLE; RELIGION GROUPING
cidi <-
  cidi_raw %>%
  filter(!is.na(de40)) %>%
  # religion
  mutate(de32_1 = ifelse(is.na(de32_1), 0, 1),
         de32_2 = ifelse(is.na(de32_2), 0, 1),
         de32_9 = ifelse(is.na(de32_9), 0, 1),
         de32_98 = ifelse(is.na(de32_98), 0, 1),
         de32_99 = ifelse(is.na(de32_99), 0, 1),
         de32_other = ifelse(de32_3|de32_4|de32_5|de32_6|de32_7|de32_8 == 1, 1, 0),
         de32_other = ifelse(is.na(de32_other), 0, 1),
         de34 = as.factor(ifelse(de34 == 8 | de34 == 9, 0, de34))) %>%
  # mental conditions
  mutate(affected_cher = ifelse(!is.na(dm40), dm40, de40),
         anx = ifelse(DSM_AGO == 1|DSM_SO == 1|DSM_GAD == 1|DSM_PDS == 1, 1, 0),
         ied = ifelse(DSM_IEDH == 1, 1, 0),
         aff = ifelse(DSM_MDE == 1|DSM_DYS == 1, 1, 0),
         alc = ifelse(DSM_ALA == 1|DSM_ALD == 1, 1, 0),
         any = ifelse(anx == 1|ied == 1|aff == 1|alc == 1, 1, 0),
         oblast = as.factor(io15)) %>%
  # 8 region variable
  mutate(region8 = derivedFactor(East = io15 == 7|io15 == 13,
                                 Eastcentral = io15 == 21|io15 == 10|io15 == 6,
                                 Krym = io15 == 1,
                                 South = io15 == 22|io15 == 15|io15 == 16,
                                 Northcentral = io15 == 26|io15 == 19|io15 == 17|io15 == 24|io15 == 12|io15 == 2|io15 == 3,
                                 Westcentral = io15 == 23|io15 == 8|io15 == 4|io15 == 18|io15 == 5,
                                 West = io15 == 11|io15 == 14|io15 == 20,
                                 Southwest = io15 == 25|io15 == 9)) %>%
  select(sampleid, affected_cher, anx, ied, aff, alc, any, 
         de32_1, de32_2, de32_other, de32_98, de32_99, oblast, region8,
         secu, strata, weight2)
```

```{r}
# Dataset with Part 2 weights on
DesignPt2 <-
  svydesign(ids = ~secu, strata = ~strata, weights = ~weight2, nest = T, 
            data = subset(cidi, cidi$weight2>0))
options(survey.lonely.psu = "adjust")
```

# Prevalence of Disorders by 8 Regions Grouping

While the 8-region grouping in the Barrington and Herron paper makes sense and is explained well, I am concerned that this creates insufficient sample sizes for those with mental disorders in each region.

```{r}
# anx
# svyglm(anx~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~anx + region8, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# alc
# svyglm(alc~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~alc + region8, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# aff
# svyglm(aff~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~aff + region8, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# ied
# svyglm(ied~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~ied + region8, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# any
# svyglm(any~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~any + region8, design = DesignPt2) %>%
  addmargins() %>%
  round()
```


# Prevalence of Disorders By Oblast

```{r}
# anx
# svyglm(anx~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~anx + oblast, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# alc
# svyglm(alc~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~alc + oblast, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# aff
# svyglm(aff~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~aff + oblast, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# ied
# svyglm(ied~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~ied + oblast, design = DesignPt2) %>%
  addmargins %>%
  round()
```

```{r}
# any
# svyglm(any~oblast, design = DesignPt2, family = quasibinomial()) %>%
#   summary
svytable(~any + oblast, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

# Frequency Counts of Mental Disorders

```{r}
# mental disorder frequency counts
svytable(~anx, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~aff, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~alc, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~ied, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~any, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

# Mental Disorders by Religion

```{r}
# anxiety and religion
svytable(~anx + de32_1, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~anx + de32_2, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~anx + de32_other, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

```{r}
# affective and religion
svytable(~aff + de32_1, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~aff + de32_2, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~aff + de32_other, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

```{r}
# alcohol and religion
svytable(~alc + de32_1, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~alc + de32_2, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~alc + de32_other, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

```{r}
# ied and religion
svytable(~ied + de32_1, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~ied + de32_2, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~ied + de32_other, design = DesignPt2) %>%
  addmargins() %>%
  round()
```

```{r}
# any and religion
svytable(~any + de32_1, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~any + de32_2, design = DesignPt2) %>%
  addmargins() %>%
  round()
svytable(~any + de32_other, design = DesignPt2) %>%
  addmargins() %>%
  round()
```












